---
title: "PA1_template"
author: "Zhengmao Zhu"
date: "August 21, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Introduction
It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a Fitbit, Nike Fuelband, or Jawbone Up. These type of devices are part of the “quantified self” movement – a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

This assignment makes use of data from a personal activity monitoring device. This device collects data at 5 minute intervals through out the day. The data consists of two months of data from an anonymous individual collected during the months of October and November, 2012 and include the number of steps taken in 5 minute intervals each day.

The data for this assignment can be downloaded from the course web site:

- Dataset: [Activity monitoring data](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)

The variables included in this dataset are:

steps: Number of steps taking in a 5-minute interval (missing values are coded as 𝙽𝙰) 
date: The date on which the measurement was taken in YYYY-MM-DD format 
interval: Identifier for the 5-minute interval in which measurement was taken 
The dataset is stored in a comma-separated-value (CSV) file and there are a total of 17,568 observations in this dataset.

#Code for reading in the dataset and/or processing the data
```{r load & process}
raw_data=read.csv("C:\\Users\\frozenl\\Desktop\\coursera\\project 7\\activity.csv",header = TRUE, sep=",",na.strings = "NA")
summary(raw_data)
str(raw_data)
head(raw_data)
raw_data$date=as.Date(raw_data$date,format="%Y-%m-%d")
raw_data$interval=factor(raw_data$interval)
```

#Histogram of the total number of steps taken each day
```{r hist of each day steps}
data=raw_data[!(is.na(raw_data$steps)), ]
day_steps=aggregate(steps~date, data=data, sum)
names(day_steps)=c("date","steps")
hist(day_steps$steps, breaks=10, col="blue", xlab="Number of Steps", main="Histogram of the total number of steps taken each day")
```

#Mean and median number of steps taken each day
```{r mean & median}
mean(day_steps$steps)
median(day_steps$steps)
```

#Time series plot of the average number of steps taken
```{r time series plot of ave steps}
interval_steps=aggregate(data$steps, by=list(interval=data$interval),FUN=mean)
names(interval_steps)=c("interval","ave_steps")
plot(levels(interval_steps$interval), interval_steps$ave_steps,type="l",col="blue",
     xlab="Interval", ylab="Average Number of Steps", main="Average Daily Activity")
```

#The 5-minute interval that, on average, contains the maximum number of steps
```{r maxiumum number of steps}
max_steps=max(interval_steps$ave_steps)
max_steps
max_interval=interval_steps$interval[which.max(interval_steps$ave_steps)]
max_interval
```

#Code to describe and show a strategy for imputing missing data
```{r filling missing data}
sum(is.na(raw_data))
sum(is.na(raw_data$steps))
sum(is.na(raw_data$date))
sum(is.na(raw_data$interval))
processed_data=raw_data
position=which(is.na(raw_data$steps))
processed_data[position, ]$steps=unlist(lapply(position, FUN=function(position){
    interval_steps[raw_data[position, ]$interval==interval_steps$interval, ]$ave_steps
}))
summary(processed_data)
```

#Histogram of the total number of steps taken each day after missing values are imputed
```{r hist of each day steps with filling missing data}
processed_day_steps=aggregate(steps~date, data=processed_data, sum)
names(processed_day_steps)=c("date","steps")
hist(processed_day_steps$steps, breaks=10, col="blue", xlab="Number of Steps", main="Histogram of the total number of steps taken each day")
```

#Mean and median number of steps taken each day
```{r mean & median with filling missing data}
mean(processed_day_steps$steps)
median(processed_day_steps$steps)
```

From the results, we know that compared with raw data, the mean of filling data is unchanged and the median of the filling data changes from 10765 to 10766.19, which is same as the mean of the filling data.

#Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends
```{r compare steps of weekdays and weekend}
week_days=c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday")
processed_data$day=as.factor(ifelse(is.element(weekdays(as.Date(processed_data$date)),week_days), "Weekday", "Weekend"))
week_steps=aggregate(steps~interval + day, data=processed_data, mean)
library(lattice)
xyplot(week_steps$steps ~ week_steps$interval | week_steps$day, layout = c(1,2), type ="l", ylab="Number of Steps", xlab="Interval", main="Average Steps per Day by Interval")
```
